Search, Compress, Compile: Library Learning in Neurally-Guided Bayesian Program Learning

### 2018-06-07 nips_2018.tex
fixed typos
nampi
almost 4 pages

\begin{table}[t!]
  \makebox[\textwidth][c]{
    \scriptsize
  \tabcolsep=3pt
  \renewcommand\code\texttt
  \renewcommand\codechar[1]{\texttt{"#1"}}
  \newcommand{\helpSize}{0.25cm}
  \begin{tabular}{>{\hspace{-0em}}c<{\hspace{-1em}}>{\hspace{-1em}}c<{\hspace{-1em}}>{\hspace{-2.5em}}c<{\hspace{-0.8em}}>{\hspace{-1em}}c<{\hspace{-1em}}}
    \toprule
    &{\normalsize List Functions}&{\normalsize Text Editing}&{\normalsize Symbolic Regression}\\\midrule
    \rotatebox[origin=c]{90}{\normalsize \pop{Programs} \& Tasks}&{\tabcolsep=4pt
      \begin{tabular}{cc}
        \begin{tabular}{c}
          \code{[7\, 0\, 2]}$\to $\code{[7\, 0\, 2\, 9]}         \\
          \code{[9\, 3]}$\to $\code{[9\, 3\, 9]} \\
          \pop{\code{$f(\ell) = $}\code{($f_1$ $\ell$ 9)}}       \\
          \\
          \\
          \code{[2\, 7\, 8\, 1]}$\to $\code{8}               \\
          \hspace{0.15cm}\code{[3\, 19\, 14]}$\to $\code{19}                \\
          \pop{\code{$f(\ell) = $}\code{($f_2$ $\ell$)}}
        \end{tabular}
        &
        \hspace{-0.3cm}\begin{tabular}{c}
          \code{[7\, 3]}$\to $\code{False}                              \\
          \hspace{0.3cm}\code{[3]}$\to $\code{False}                    \\
          \hspace{-0.3cm}\code{[9\, 0\, 0]}$\to $\code{True\phantom{e}} \\
          \hspace{0.3cm}\code{[0]}$\to $\code{True\phantom{e}}                        \\
          \hspace{-0.3cm}\code{[0\, 7\, 3]}$\to $\code{True\phantom{e}}                \\
          \pop{\code{$f(\ell) = $}\code{($f_3$ $\ell$ 0)}}
        \end{tabular}
      \end{tabular}
    }
    &
    \hspace{-0.5cm}\begin{tabular}{c}
      +106 769-438$\to $106.769.438\\%&Nancy FreeHafer $\longrightarrow$ Dr. Nancy\\
      +83 973-831$\to $83.973.831\\
      \pop{$f(\text{\code{s}}) = $\code{(}$f_0$\code{  \codechar{.} \codechar{,}
      }}\\
      \hspace{1.25cm}\pop{\code{($f_0$ \codechar{.} \codechar{ }}}\\
      \hspace{1.4cm}\pop{\code{(cdr s)))}}\\
      ~\\
      Temple Anna H $\to $TAH\\
      Lara Gregori$\to $LG\\
      \pop{$f(\text{\code{s}}) = $\code{(}$f_2$\code{ s)}}\\
    \end{tabular}
    &
    \begin{tabular}{cc}
      \includegraphics[width = 3em]{figures/functions/4.png}&
      \includegraphics[width = 3em]{figures/functions/146}\\
      \pop{\code{$f($x$) = $($f_1$ x)}}&    \pop{\code{$f($x$) = $($f_6$ x)}}\\
      ~\\
      \includegraphics[width = 3em]{figures/functions/112.png}&
        \includegraphics[width = 3em]{figures/functions/92.png}
      \\
      \pop{\code{$f($x$) = $($f_4$ x)}}&    \pop{\code{$f($x$) = $($f_3$ x)}}\\

    \end{tabular}
    ~\\
    \midrule
    \rotatebox[origin=c]{90}{\normalsize \popp{DSL}}&
    \hspace{0cm}\begin{tabular}{l}
      % $f_0($\code{r}$,\ell) \,=\, $\code{(foldr $\ell$ r ($\lambda$ (x a)}\\
      % \phantom{$f_0($\code{r}$,\ell) \,=\, $\code{(foldr}}\code{(cons (index (length a) $\ell$) a)))}\\
      % \hspace{\helpSize}($f_1$: \emph{Get the largest number})\\
      % $f_0(\ell) \,=\, $\code{(foldr $\ell$ 0 ($\lambda$ (x a)}\\\hspace{0.8cm}\code{ (if (> a x) a x)))}\\
      % \hspace{\helpSize}($f_1$: \emph{Get the largest number in $\ell$})\\
      %NOTE: this is the actual invention, but I removed a redundant lambda
      % below: $f_0(\ell,$\code{r}$) \,=\, $\code{(foldr r $\ell$ ($\lambda$ (x a) (cons x a)))}\\
      \popp{$f_0(\ell,$\code{r}$) \,=\, $\code{(foldr r $\ell$ cons)}}\\
      \hspace{\helpSize}($f_0$: \emph{Append lists }\code{r}\emph{ and  $\ell$})\\
      \popp{$f_1(\ell,$\code{k}$) \,=\, $\code{($f_0$ (cons k nil) $\ell$)}}\\
      \hspace{\helpSize}($f_1$: \emph{Append the number }\code{k}\emph{ to $\ell$})\\
      %(lambda (fold $0 0 (lambda (lambda (if (gt? $0 $1) $0 $1)))))
      \popp{$f_2(\ell) \,=\, $\code{(foldr $\ell$ 0 ($\lambda$ (x a)}}\\
      \popp{\phantom{$f_2(\ell) \,=\, $}\code{(if (> a x) a x)))}}\\
      \hspace{\helpSize}($f_2$: \emph{Maximum element in list $\ell$})\\
      \popp{$f_3(\ell,$\code{k}$) \,=\, $\code{(foldr $\ell$ (is-nil $\ell$)}}\\
      \phantom{$f_1(\ell,$}
      \popp{\code{($\lambda$ (x a) (if a a (= k x))))}}\\
      \hspace{\helpSize}($f_2$: \emph{Whether $\ell$ contains }\code{k})\\
    \end{tabular}&

  \hspace{0.5cm}\begin{tabular}{l}
    \popp{$f_0($\code{s}$,$\code{a}$,$\code{b}$) \,=\, $\code{(map ($\lambda$
    (x)}}\\
    \popp{\hspace{1cm}\code{ (if (= x a) b x)) s)}}\\
      \hspace{\helpSize}($f_0$: \emph{Performs character substitution)}\\
      \popp{$f_1($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s ($\lambda$ (x
      a)}\\\hspace{1.1cm}\code{ (cdr (if (= c x) s a))))}}\\
        \hspace{\helpSize}($f_1$: \emph{Drop characters from }\code{s}\emph{ until  }\code{c}\emph{ reached})\\
      \popp{$f_2($\code{s}$) \,=\, $\code{(unfold s is-nil car
      }}\\
      \popp{\hspace{1cm}\code{($\lambda$ (z) (}$f_1$\code{ z \codechar{ })))}}\\
        \hspace{\helpSize}($f_2$: \emph{Abbreviates a sequence of words})\\
        \popp{$f_3($\code{a}$,$\code{b}$) \,=\, $\code{(foldr a b cons)}}\\
      \hspace{\helpSize}($f_3$: \emph{Concatenate strings }\code{a}\emph{ and }\code{b})
  \end{tabular}&

@@

  \end{tabular}
  \\\bottomrule\\
\end{tabular}}
\caption{Top: Tasks from each domain, each followed by the programs \system discovers for them. Bottom: Several examples from learned DSL. Notice that learned DSL primitives can call each other.}\label{initialExampleDSL}
\end{table}

@@

\system takes as input a set of \emph{tasks}, written $X$, each of which is a program synthesis problem.
It has at its disposal a domain-specific \emph{likelihood model}, written $\probability[x|p]$, which scores the likelihood of a task $x\in X$ given a program $p$.%% \footnote{For example, for string editing,
%%   the likelihood is 1 if the program predicts the observed outputs on the observed inputs,
%% and 0 otherwise.}
Its goal is to solve each of the tasks by writing a program,
and also to infer a DSL, written $\mathcal{D}$.
We equip $\mathcal{D}$ with a real-valued weight vector $\theta$, and together
$(\mathcal{D},\theta)$ define a generative model over programs.
We frame our goal as maximum a posteriori (MAP) inference of $(\mathcal{D},\theta)$ given $X$.
Writing $J$ for the joint probability of $(\mathcal{D},\theta)$ and $X$, we want the $\mathcal{D}^*$ and $\theta^*$ solving:
\begin{align}\label{intractableObjectives}
\nonumber  J(\mathcal{D},\theta)\triangleq \probability[\mathcal{D},\theta]\prod_{x\in X} \sum_p \probability[x|p]\probability[p|\mathcal{D},\theta]\\
  \mathcal{D}^* = \argmax_{\mathcal{D}}\int J(\mathcal{D},\theta)\;\mathrm{d}\theta \qquad
  \theta^* =\argmax_\theta J(\mathcal{D}^*,\theta)
\end{align}

>>>>

\begin{table}[t!]
  \makebox[\textwidth][c]{
    \scriptsize
  \tabcolsep=3pt
  \renewcommand\code\texttt
  \renewcommand\codechar[1]{\texttt{"#1"}}
  \newcommand{\helpSize}{0.25cm}
  \begin{tabular}{>{\hspace{-0em}}c<{\hspace{-1em}}>{\hspace{-1em}}c<{\hspace{-1em}}>{\hspace{-2.5em}}c<{\hspace{-0.8em}}>{\hspace{-1em}}c<{\hspace{-1em}}}
    \toprule
    &{\normalsize List Functions}&{\normalsize Text Editing}&{\normalsize Symbolic Regression}\\\midrule
    \rotatebox[origin=c]{90}{\normalsize \pop{Programs} \& Tasks}&{\tabcolsep=4pt
      \begin{tabular}{cc}
        \begin{tabular}{c}
          \code{[7\, 2\, 3]}$\to $\code{[7\, 3]}         \\
          \code{[1\, 2\, 3\, 4]}$\to $\code{[3\, 4]} \\
          \code{[4\, 3\, 2\, 1]}$\to $\code{[4\, 3]} \\
          \pop{\code{$f(\ell) = $}\code{($f_1$ $\ell$ ($\lambda$ (x)}}\\
          \hspace{1.15cm}\pop{\code{(> x 2)))}}       \\
          \\
          \\
          \code{[2\, 7\, 8\, 1]}$\to $\code{8}               \\
          \hspace{0.15cm}\code{[3\, 19\, 14]}$\to $\code{19}                \\
          \pop{\code{$f(\ell) = $}\code{($f_2$ $\ell$)}}
        \end{tabular}
        &
        \hspace{-0.3cm}\begin{tabular}{c}
          \code{[7\, 3]}$\to $\code{False}                              \\
          \hspace{0.3cm}\code{[3]}$\to $\code{False}                    \\
          \hspace{-0.3cm}\code{[9\, 0\, 0]}$\to $\code{True\phantom{e}} \\
          \hspace{0.3cm}\code{[0]}$\to $\code{True\phantom{e}}                        \\
          \hspace{-0.3cm}\code{[0\, 7\, 3]}$\to $\code{True\phantom{e}}                \\
          \pop{\code{$f(\ell) = $}\code{($f_3$ $\ell$ 0)}}
        \end{tabular}
      \end{tabular}
    }
    &
    \hspace{-0.5cm}\begin{tabular}{c}
      +106 769-438$\to $106.769.438\\%&Nancy FreeHafer $\longrightarrow$ Dr. Nancy\\
      +83 973-831$\to $83.973.831\\
      \pop{$f(\text{\code{s}}) = $\code{(}$f_0$\code{  \codechar{.} \codechar{-}
      }}\\
      \hspace{1.25cm}\pop{\code{($f_0$ \codechar{.} \codechar{ }}}\\
      \hspace{1.4cm}\pop{\code{(cdr s)))}}\\
      ~\\
      Temple Anna H $\to $TAH\\
      Lara Gregori$\to $LG\\
      \pop{$f(\text{\code{s}}) = $\code{(}$f_2$\code{ s)}}\\
    \end{tabular}
    &
    \begin{tabular}{cc}
      \includegraphics[width = 3em]{figures/functions/4.png}&
      \includegraphics[width = 3em]{figures/functions/146}\\
      \pop{\code{$f($x$) = $($f_1$ x)}}&    \pop{\code{$f($x$) = $($f_6$ x)}}\\
      ~\\
      \includegraphics[width = 3em]{figures/functions/112.png}&
        \includegraphics[width = 3em]{figures/functions/92.png}
      \\
      \pop{\code{$f($x$) = $($f_4$ x)}}&    \pop{\code{$f($x$) = $($f_3$ x)}}\\

    \end{tabular}
    ~\\
    \midrule
    \rotatebox[origin=c]{90}{\normalsize \popp{DSL}}&
    \hspace{0cm}\begin{tabular}{l}
      % $f_0($\code{r}$,\ell) \,=\, $\code{(foldr $\ell$ r ($\lambda$ (x a)}\\
      % \phantom{$f_0($\code{r}$,\ell) \,=\, $\code{(foldr}}\code{(cons (index (length a) $\ell$) a)))}\\
      % \hspace{\helpSize}($f_1$: \emph{Get the largest number})\\
      % $f_0(\ell) \,=\, $\code{(foldr $\ell$ 0 ($\lambda$ (x a)}\\\hspace{0.8cm}\code{ (if (> a x) a x)))}\\
      % \hspace{\helpSize}($f_1$: \emph{Get the largest number in $\ell$})\\
      %NOTE: this is the actual invention, but I removed a redundant lambda
      % below: $f_0(\ell,$\code{r}$) \,=\, $\code{(foldr r $\ell$ ($\lambda$ (x a) (cons x a)))}\\
      \popp{$f_0(\ell,$\code{r}$) \,=\, $\code{(foldr r $\ell$ cons)}}\\
      \hspace{\helpSize}($f_0$: \emph{Append lists }\code{r}\emph{ and  $\ell$})\\
      \popp{$f_1(\ell,$\code{p}$) \,=\, $\code{(foldr $\ell$ nil ($\lambda$ (x a)}}\\
      \hspace{0.5cm}\popp{\code{(if (p x) (cons x a) a)))}}\\
      \hspace{\helpSize}($f_1$: \emph{Higher-order filter function})\\
      %(lambda (fold $0 0 (lambda (lambda (if (gt? $0 $1) $0 $1)))))
      \popp{$f_2(\ell) \,=\, $\code{(foldr $\ell$ 0 ($\lambda$ (x a)}}\\
      \popp{\phantom{$f_2(\ell) \,=\, $}\code{(if (> a x) a x)))}}\\
      \hspace{\helpSize}($f_2$: \emph{Maximum element in list $\ell$})\\
      \popp{$f_3(\ell,$\code{k}$) \,=\, $\code{(foldr $\ell$ (is-nil $\ell$)}}\\
      \phantom{$f_1(\ell,$}
      \popp{\code{($\lambda$ (x a) (if a a (= k x))))}}\\
      \hspace{\helpSize}($f_2$: \emph{Whether $\ell$ contains }\code{k})\\
    \end{tabular}&

  \hspace{0.5cm}\begin{tabular}{l}
    \popp{$f_0($\code{s}$,$\code{a}$,$\code{b}$) \,=\, $\code{(map ($\lambda$
    (x)}}\\
    \popp{\hspace{1cm}\code{ (if (= x a) b x)) s)}}\\
      \hspace{\helpSize}($f_0$: \emph{Performs character substitution)}\\
      \popp{$f_1($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s ($\lambda$ (x
      a)}\\\hspace{1.1cm}\popp{\code{ (cdr (if (= c x) s a))))}}}\\
        \hspace{\helpSize}($f_1$: \emph{Drop characters from }\code{s}\emph{ until  }\code{c}\emph{ reached})\\
      \popp{$f_2($\code{s}$) \,=\, $\code{(unfold s is-nil car
      }}\\
      \popp{\hspace{1cm}\code{($\lambda$ (z) (}$f_1$\code{ z \codechar{ })))}}\\
        \hspace{\helpSize}($f_2$: \emph{Abbreviates a sequence of words})\\
        \popp{$f_3($\code{a}$,$\code{b}$) \,=\, $\code{(foldr a b cons)}}\\
      \hspace{\helpSize}($f_3$: \emph{Concatenate strings }\code{a}\emph{ and }\code{b})
  \end{tabular}&

@@

  \end{tabular}
  \\\bottomrule\\
\end{tabular}}
\caption{Top: Tasks from each domain, each followed by the programs \system discovers for them. Bottom: Several examples from learned DSL. Notice that learned DSL primitives can call each other, and that \system rediscovers higher-order functions like \code{filter} ($f_1$ in List Functions)}\label{initialExampleDSL}
\end{table}

@@

\system takes as input a set of \emph{tasks}, written $X$, each of which is a program synthesis problem.
It has at its disposal a domain-specific \emph{likelihood model}, written $\probability[x|p]$, which scores the likelihood of a task $x\in X$ given a program $p$. %% \footnote{For example, for string editing,
%%   the likelihood is 1 if the program predicts the observed outputs on the observed inputs,
%% and 0 otherwise.}
Its goal is to solve each of the tasks by writing a program,
and also to infer a DSL, written $\mathcal{D}$.
We equip $\mathcal{D}$ with a real-valued weight vector $\theta$, and together
$(\mathcal{D},\theta)$ define a generative model over programs.
We frame our goal as maximum a posteriori (MAP) inference of $(\mathcal{D},\theta)$ given $X$.
Writing $J$ for the joint probability of $(\mathcal{D},\theta)$ and $X$, we want the $\mathcal{D}^*$ and $\theta^*$ solving:
\begin{align}\label{intractableObjectives}
\nonumber  J(\mathcal{D},\theta)\triangleq \probability[\mathcal{D},\theta]\prod_{x\in X} \sum_p \probability[x|p]\probability[p|\mathcal{D},\theta]\\
  \mathcal{D}^* = \argmax_{\mathcal{D}}\int J(\mathcal{D},\theta)\;\mathrm{d}\theta \qquad
  \theta^* =\argmax_\theta J(\mathcal{D}^*,\theta)
\end{align}
